---
outpur_dir: "."
title: "Inserting R Code in Markdown"
author: "Richard Goodman"
date: "11 December 2020"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

\ 

# Importing and manipulating Data

First make sure there is a folder called data in your wd

Let us download the `gapminder_data.csv` dataset into our project

```{r download-file, message = FALSE}

download.file(url="https://raw.githubusercontent.com/cambiotraining/reproducibility-training/master/data/gapminder_data.csv", destfile="data/gapminder_data.csv")

```

Start to load the dataset 

```{r load-data, message=FALSE, warning = FALSE}

# Load all libraries
library(tidyverse) # use for data manipulation and visualisation
library(kableExtra) # used for kbl()
library(rmarkdown) # used for paged_table function
library(ggpubr) #used for ggarrange function

# read file into R 
pop_data = read_csv("data/gapminder_data.csv")

# create a table with data from Euro countries in 2007 - ordered by life expectancy 
euro_data_tbl = pop_data %>% 
  # Filter by continent and year 
  filter(continent == "Europe" & year == 2007) %>% 
  # remove continent and year (since they are all the same)
  select(-continent, -year) %>% 
  # arrange by life expecntancy 
  arrange(desc(lifeExp)) %>% 
  # Rename columns names to how we want them 
  rename(Country = country, 
         "Population size" = pop, 
         "Life Expectancy" = lifeExp,
         "GDP" = gdpPercap)





```

\ 

# Tables in R Markdown

We will use the `kableExtra` package to design some tables. The vignette for kableExtra can be found [here](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
)

The results in euro_data_tbl are displayed in the table below:


```{r Making tables with kable}

euro_data_tbl %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", full_width = F)) %>% 
                  scroll_box(width = "100%", height = "200px" )

```

A better way to display long tables is using the function `paged_table()` in the `rmarkdown` library can be used to do this

```{r paged-table}

paged_table(euro_data_tbl)

```

\ 

# Adding images

Adding images is straightforward and does not require using a specific R Markdown function.

Create a new dataset `euro_data_fig` by filtering the `pop_data` tibble to contain only data from Europe. Draw a plot to display the `lifeExp` on the y axis and `year` on the x axis. Use `geom_violin()` to draw this as a violin plot to show the distrubution of the data across each year and save it in a `1euro_plot` variable.

```{r adding ggplot images}

euro_plot = pop_data %>% 
  filter(continent == "Europe") %>% 
  select(-continent) %>% 
  # Add factor() to ensure year is treated as discrete 
  ggplot(aes(x = factor(year), y = lifeExp)) +
  # use geom_violin
  geom_violin() +
  # plot the median as a point on the violin 
  stat_summary(fun = median, geom = "point") +
  ylim(40, 85)

# display plot 
euro_plot

```

 

Create a new dataset `uk_data_fig` by filtering the `pop_data` tibble to contain only data from the United Kingdom. Draw a scatter plot to display the `lifeExp` on the y axis and `year` on the x axis and save it in a `uk_plot` variable.
Draw the `euro_plot` created in the previous challenge next to a `uk_plot` using the `ggarrange()` function. Label the plots A and B respectively.

```{r adding multiple-plots }

uk_plot = pop_data %>% 
  # filter for United Kingdom 
  filter(country == "United Kingdom") %>%
  # use mutate to convert year to factor 
  mutate(year = as_factor(year)) %>% 
  # Plot a scatterplot with geom_point
  ggplot(aes(x = year, y = lifeExp)) +
  geom_point() +
  ylim(40, 85)

# display plot 
uk_plot

```

ggarrage() is a function present in the ggpubr pacakge that can place two or more figures next to each other

```{r}

# Use ggarrange to arrange the uk_plot and euro_plot 
ggarrange(uk_plot, euro_plot, ncol = 2, nrow = 1, labels = c("A", "B"))

```


`rmarkdown::render("inserting-code-in-rmarkdown.Rmd")`
`rmarkdown::render("index.Rmd")`




